As scientists, we also have the opportunity to intellectual impact of our work can be made tangible through technology.
Although my interest lies in fundamental research, by probing the algorithmic principles of biology I hope to contribute to the development of more capable and versatile artificial intelligence systems.
Beneath my own curiosity and ambition, I feel a moral imperative to participate in this work.
When reflecting on the potential human impact of advances in computing, I am reminded of my father, who volunteers with the Dial-A-Bus program in Benton county.
He serves individuals who would be otherwise unable to get around town, curtailing their ability to fully participate in the community.
I hope that, through applications that counteract disability and free us from dangerous or simply menial tasks, stronger artificial intelligence will enable people to more fully exercise their human capabilities and, thereby, lay a fuller claim to their humanity.
I see glimmers of progress towards this goal in the work I've done with the Swarm Lab, the Fowler lab, and, now, the Ofria lab.

To translate my research into applied advances in artificial intelligence, I will work in close collaboration with industry.
I participated in such work during my my time with the Finn small fruits breeding laboratory at the USDA.
Pushing our berry cultivars out into the fresh fruit and processed food markets requires extensive and deliberate legwork.
We hosted regular symposiums around brief and actionable presentations on best practices and made a point to always wrap up with taste-tests berry sauce with ice cream.
I learned how to develop personal relationships, how to concretely explain --- and demonstrate --- a compelling use case for my work in the context of \textit{their practices}.
The scale our agricultural partners can achieve made our work at the USDA even more rewarding;
we get to see our cultivars fill vast fields that sweep over the horizon and, ultimately, in muffins, ice cream, farmers' markets, and roadside stands.
From my vantage, I see a wave of bio-inspired computing techniques sweeping through industry.
I am ready to work with industrial partners to add my research to that wave.
